Case for Support

### 2021-11-17 application/Je-S_sections.tex

The overall goal of this proposal is to extend, enhance, maintain and support Bonsai, a fully integrated software environment to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models.

@@

1. Develop, test and deploy a new software workflow to incorporate external packages into Bonsai's C# environment. \textcolor{red}{I believe this is already done. All existing bonsai community packages already depend on the existence of the extendable plugin architecture -- perhaps you mean "languages" rather than "packages"? i.e. MATLAB, Python, R, in that case I would rewrite as below}

1. Develop, test and deploy new foreign language interfaces to allow extending the Bonsai environment with new packages developed in languages other than C#, namely MATLAB, Python and R.

>>>>

The overall goal of this proposal is to extend, enhance, maintain and support Bonsai, a fully integrated software environment enabling cutting-edge reproducible neuroscience and behavioural science experiments, with machine-intelligence-enabled, real-time neuroinformatics methods.

@@

1. Develop, test and deploy new programming language interfaces to allow extending the Bonsai environment with new packages developed in languages other than C#, namely MATLAB, Python and R.

### 2021-11-17 application/backgroundToTheResource.tex

Progress is thus held back by the challenges of designing and implementing new experimental protocols, and of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs.
%
\textbf{Reproducibility}, in particular, requires standardised tools for experimental control and analysis, shared between
laboratories \citep{baker16,ioannidis05}. Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are scarce in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).

>>>>

Progress is thus held back by the challenges of designing and implementing new experimental protocols, and of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs.
%
\textbf{Reproducibility}, in particular, requires standardised tools for experimental control and analysis, shared between
laboratories \citep{baker16,ioannidis05}. Although such tools exist in genetics, astronomy, physics and medicine \citep{fishEtAl16,abdallaEtAl18,cern18,dickinsonEtAl16,bycroftEtAl18}, they are scarce in neuroscience, with documented consequences for the reliability of ``discoveries'' \citep{baker16,botvinikNezerEtAl20,buttonEtAl13}.

### 2021-11-17 application/bonsai.tex

Bonsai has been in adopted in hundreds of laboratories worldwide and has the largest user base in the systems neuroscience community (Figure~\ref{fig:bonsai}D-F).  In the last year alone, more than 1,000 new users incorporated Bonsai into their experimental protocols. The surprising rate of adoption of Bonsai in non-programmer experimental
labs highlights the need for accessible design tools that enable
state-of-the-art technology but also allow researchers to stay in control and flexibly
change their experimental paradigms.  Many open-source software tools are
either inaccessible to non-programmers, or too constrained to be of general use
outside their narrow domain of application. Bonsai has been successful because
it has straddled this gap.

The language has also helped to potentiate the growing wave of foundational
open hardware initiatives, such as the OpenEphys \citep{siegleEtAl17} and UCLA
Miniscope~\citep{caiEtAl16}, making it possible to quickly combine and
integrate these tools into new experiments~\citep{buccinoEtAl18}.
%
Bonsai has been adopted in large neuroscience undertakings like the
International Brain
Laboratory\footnote{https://www.internationalbrainlab.com/}
and the Allen Institute for Brain
Science\footnote{https://alleninstitute.org/what-we-do/brain-science/}.
%
Key to the wide adoption is the reproducibility that Bonsai ensures.  Data acquisition and experimental control protocols can be 
replicated in any laboratory by just sharing a Bonsai configuration file. 
%
% LEAVE OUT FOR THIS GRANT:
% More recently, Bonsai has started to expand outside the domain of neuroscience
% into biomedical research and biotechnology tool development, and even outside
% academia into public outreach and education programs.

>>>>

Bonsai has been adopted in hundreds of laboratories worldwide and has the largest user base in the systems neuroscience community (Figure~\ref{fig:bonsai}D-F).  In the last year alone, more than 1,000 new users incorporated Bonsai into their experimental protocols. The rapid rate of adoption of Bonsai in non-programmer experimental
labs highlights the need for accessible design tools that enable
state-of-the-art technology but also allow researchers to stay in control and flexibly
change their experimental paradigms.  Many open-source software tools are
either inaccessible to non-programmers, or too constrained to be of general use
outside their narrow domain of application. Bonsai has been successful because
it bridges this gap.

Bonsai also supports the growing wave of foundational
open hardware initiatives, such as OpenEphys \citep{siegleEtAl17} and UCLA
Miniscope~\citep{caiEtAl16}, allowing these tools to be quickly combined and integrated into new experiments~\citep{buccinoEtAl18}.
%
Bonsai has been adopted in large neuroscience undertakings like the
International Brain
Laboratory\footnote{https://www.internationalbrainlab.com/}
and the Allen Institute for Brain
Science\footnote{https://alleninstitute.org/what-we-do/brain-science/}.
%
Key to the wide adoption is the reproducibility that Bonsai ensures.  Data acquisition and experimental control protocols can be 
replicated in any laboratory just by sharing a Bonsai configuration file. 
%
% LEAVE OUT FOR THIS GRANT:
% More recently, Bonsai has started to expand outside the domain of neuroscience
% into biomedical research and biotechnology tool development, and even outside
% academia into public outreach and education programs.

### 2021-11-17 application/resourceManagementandSustainability.tex

\section{Resource management}
%
As an open source project developed over almost 10 years, Bonsai 
has well-established management practices built around the user community.
%
The source code for the core system and the majority of extension packages
are hosted at \url{github.com/bonsai-rx}.
%
The documentation, including learning materials and all relevant community
links are on a publicly accessible website at \url{bonsai-rx.org}.
%
A community user forum is a public Google Group with more than 700 active users (Figure~\ref{fig:bonsai}f) and several
new questions and answers exchanged every day.
%
Package installation and updates are supported by standalone package
management software included within the core Bonsai distribution and hosted
in a shared centralized and curated repository.
%
This ensures that bugfixes and enhancements are easily available but
also importantly allows for reproducible deployment of user environments
to ensure working experiments are not disrupted by changes in the system.
Large-scale experimental projects such as the International Brain Laboratory
have leveraged this capability of Bonsai to standardize data acquisition and
control software across dozens of laboratories around the world.
%
Community technology previews are made available through the package manager,
allowing early adopters to opt-in to the latest features
before they are generally available. This allows the community
to help report issues with new features on GitHub, thereby
distributing bug reporting and triage and overall improving the quality of
stable releases.

>>>>

\section{Resource management}
%
As an open source project developed over almost 10 years, Bonsai 
has well-established management practices built around the user community.
%
The source code for the core system and the majority of extension packages
are hosted at \url{github.com/bonsai-rx}.
%
The documentation, including learning materials and all relevant community
links are on a publicly accessible website at \url{bonsai-rx.org}.
%
A community user forum is a public Google Group with more than 700 active users (Figure~\ref{fig:bonsai}f) and several
new questions and answers exchanged every day.
%
Package installation and updates are supported by standalone package
management software included within the core Bonsai distribution and hosted
in a shared centralised and curated repository.
%
This ensures that bugfixes and enhancements are easily available but
also importantly allows for reproducible deployment of user environments
to ensure working experiments are not disrupted by changes in the system.
Large-scale experimental projects such as the International Brain Laboratory
have leveraged this capability of Bonsai to standardise data acquisition and
control software across dozens of laboratories around the world.
%
Community technology previews are made available through the package manager,
allowing early adopters to opt-in to the latest features
before they are generally available. This allows the community
to help report issues with new features on GitHub, thereby
distributing bug reporting and triage and overall improving the quality of
stable releases.
